Importance of depth refugia for reef resilience and intervention on the Great Barrier Reef under future climate change

### 2025-10-05 paper/paper_draft.qmd
Updates to supplementary material (excluding ADRIA-CoralBlox changes pending)

The CID time series clustering metric corrects the Euclidean distance between pairs of time
series (which accounts for magnitude differences) with the complexity of each time series
(obtained through a measure of their variability) to ensure reefs are clustered based on
their underlying temporal dynamics rather than simply their average coral cover. The process
creates a distance matrix of n \* n dimensions (where n is the number of time series in the
group) that is clustered using a k-medoids clustering algorithm [cf. @steinmann2020]. The
three reef clusters are labelled with qualitative-ordered clusters from Low, Medium and High
based on their median cover relative to each other cluster over the 2030-2060 period. The
selected time frame prevents the classification from being overly influenced by any steep
declines in the initial years of simulations and avoids the post-2060 time period where
reefs are projected to be under severe stress and in a declined state [@bozec2025].

## ACCESS-ESM1-5

@@

![Scatter plot indicating reef median depth (metres) and log10 weighted incoming
connectivity for each reef. Point colours represent the median number of years each reef
maintains positive carbonate budgets across each $theta$ value from 10-20% live coral cover.
Displays results for ACCESS-ESM1-5
GCM.](../figs/ACCESS-ESM1-5/depth_conn_mean_carbonate_budget.png){#sfig-access-esm1-5-mean-carb-budget
fig-align="center" width="14.82cm"}

@@

Results are only shown and discussed for ACCESS-ESM1-5, an intermediate ECS scenario
(@sfig-ecs). Those interested in results for the other assessed GCMs are referred to Grier
and Iwanaga [-@grier2025]. Reefs in the High cluster were consistently found to have the
greatest median reef depths across bioregions (@sfig-access-esm1-5-bioregion-depth). Across
bioregions, there were no consistent patterns between reef clusters in mean DHW values
(@sfig-access-esm1-5-bioregion-dhwmean), log10-weighted incoming connectivity
(@sfig-access-esm1-5-bioregion-weightedincoming), log10 outgoing connection strength
(@sfig-access-esm1-5-bioregion-outgoingconn), log10 source-to-sink connectivity ratio
(@sfig-access-esm1-5-bioregion-logsosi) and log10 larval retention probability
(@sfig-access-esm1-5-bioregion-selfconn). Across six of the 29 bioregions reefs in the high
cover cluster have lower log10 carrying capacity values than reefs in other clusters
(@sfig-access-esm1-5-bioregion-carrying-capacity).

@@

#### Random Forest Performance

>>>>

The CID time series clustering metric corrects the Euclidean distance between pairs of time
series with their relative complexity (which is the quotient of their variability). This is
to ensure reefs are clustered based on their underlying temporal dynamics rather than
aggregating the coral cover across all timesteps. The process creates a n \* n distance
matrix (where n is the number of time series in the group) that is used as an input to a
k-medoids clustering algorithm [cf. @steinmann2020]. The three reef clusters are labelled
with qualitative-ordered labels from Low, Medium and High based on their median cover
relative to each other cluster over the 2030-2060 period. The selected time frame prevents
the classification from being overly influenced by any steep declines in the initial years
of simulations and avoids the post-2060 time period where reefs are projected to be under
severe stress and in a declined state [@bozec2025].

## ACCESS-ESM1-5 results

ACCESS-ESM1-5 was selected for GCM-level results as an intermediate severity GCM, with an
ECS value closest to the midpoint of the IPCC likely range of ECS values (3.87 °C). The
following section includes supplemental results for analyses under this GCM. A comparison of
results across GCMs can be found in the main text, and individual GCM-level results for
other GCMs can be found at Grier and Iwanaga [-@grier2025].

@@

![Scatter plot indicating reef median depth (metres) and log10 weighted incoming
connectivity for each reef. Point colours represent the median number of years each reef
maintains positive carbonate budgets across $theta$ values from 10-20% live coral cover.
Displays results for ACCESS-ESM1-5
GCM.](../figs/ACCESS-ESM1-5/depth_conn_mean_carbonate_budget.png){#sfig-access-esm1-5-mean-carb-budget
fig-align="center" width="14.82cm"}

@@

The following section includes bioregion scale clustering results, full results for other
spatial scales (management area and GBR) can be found at Grier and Iwanaga [-@grier2025].
Results are only shown and discussed for ACCESS-ESM1-5, an intermediate ECS scenario
(@sfig-ecs). Those interested in results for the other assessed GCMs are referred to Grier
and Iwanaga [-@grier2025]. Reefs in the High cluster were consistently found to have the
greatest median reef depths across bioregions (@sfig-access-esm1-5-bioregion-depth). Across
bioregions, there were no consistent patterns between reef clusters in mean DHW values
(@sfig-access-esm1-5-bioregion-dhwmean), log10-weighted incoming connectivity
(@sfig-access-esm1-5-bioregion-weightedincoming), log10 outgoing connection strength
(@sfig-access-esm1-5-bioregion-outgoingconn), log10 source-to-sink connectivity ratio
(@sfig-access-esm1-5-bioregion-logsosi) and log10 larval retention probability
(@sfig-access-esm1-5-bioregion-selfconn). Across six of the 29 bioregions reefs in the high
cover cluster have lower log10 carrying capacity values than reefs in other clusters
(@sfig-access-esm1-5-bioregion-carrying-capacity).

@@

## Random Forest Performance
